---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# bedshear

bedshear is an R package that let's you calculate bed shear stress from combined waves and tides, and wave orbital velocity.

## Installation

You can install the development version of bedshear from github using:


``` r
# install.packages("devtools")
devtools::install_github("r4ecology/rcdo", dependencies = TRUE)
```

This depends on Rcpp, so please check you can compile C++ files that rely on the math library.

## Examples

Calculating a time series of bed shear stress from wave and tide data near Stonehaven, Scotland.

```{r stonehaven_ts, message = FALSE}
# load the necessary libraries
library(bedshear)
library(dplyr)
library(ggplot2)
# calculate and plot the bed shear stress time series
stonehaven_stress <- shear_stress(bathymetry = 40,0.0002, stonehaven_ts$tidal_velocity, stonehaven_ts$tidal_direction,stonehaven_ts$wave_height,
																		stonehaven_ts$wave_period , switch = 0, stonehaven_ts$wave_direction )
stonehaven_ts %>%
	mutate(Stress = stonehaven_stress$shearmax) %>%
	ggplot(aes(Date, Stress))+
geom_line()
```



Calculating a time series of wave orbital velocities from wave data near Stonehaven, Scotland.

```{r message = FALSE}
# load libraries
library(bedshear)
library(dplyr)
library(ggplot2)
# calculate and plot a time series of wave orbital velocities
stonehaven_orbital <- orbital_velocity(bathymetry = 40, stonehaven_ts$wave_height, stonehaven_ts$wave_period, switch = 0)
stonehaven_ts %>%
mutate(Velocity = stonehaven_orbital) %>%
ggplot(aes(Date, Velocity))+
geom_line()
```


